Finding Maximum $k$-Cliques Faster using Lazy Global Domination

### 2016-04-06 paper/paper.tex
Parallel experiments

\begin{abstract}
    A clique in a graph is a set of vertices, each of which is adjacent to every other vertex in
    this set. A $k$-clique relaxes this requirement, requiring vertices to be within a distance $k$
    of each other, rather than directly adjacent. In theory, a maximum clique algorithm can easily
    be adapted to solve the maximum $k$-clique problem. We use a state of the art maximum clique
    algorithm to show that this is feasible in practice, and introduce a lazy global domination rule
    which sometimes vastly reduces the search space. We include experimental results for a range of
    real-world and benchmark graphs, and a detailed look at random graphs.
\end{abstract}

@@

A clique in a graph is a set of vertices, each of which is adjacent to every other vertex in the
set. Finding a clique of maximum size in a graph is one of the basic \NP-hard problems
\cite{Garey:1990}; applications include geometry, coding theory, computer vision and bioinformatics
\cite{Bomze:1999,Butenko:2006}. However, when analysing real-world data, a clique may be too strong
a requirement. A $k$-clique (or sometimes $n$-clique or $s$-clique) is a relaxed form of clique,
where instead of requiring each pair of vertices to be directly adjacent, we only require that they
be connected by a path of length at most $k$ \cite{Luce:1950} \footnote{In an unfortunate clash of
notation, the name ``$k$-clique problem'' is sometimes used elsewhere to refer the decision version
of the clique problem, to distinguish it from the maximum clique problem.}.  Thus a $1$-clique
\emph{is} a clique, a $2$-clique may be thought of as ``a group of people, all of whom either know
each other or have a mutual acquaintance'', and so on. We illustrate this in
\cref{figure:k-cliques}.  Determining the size of a maximum $k$-clique is \NP-hard for any fixed $k$
\cite{Bourjolly:2002}.

@@

Here we give experimental results on a range of standard benchmarks, and on real-world and random
graphs. Where timing results are reported, the experiments were run on a machine with Intel E5-2687W
v3 processors, and single-threaded runtimes are given. The time taken to read in the graph from a
file is excluded, but preprocessing time (including the construction of $G^k$ and the bitset
encoding) is included. We use the term \emph{nodes} to refer to the number of recursive calls made
by the branch-and-bound part of the algorithm.

@@

>>>>

\begin{abstract}
    A clique in a graph is a set of vertices, each of which is adjacent to every other vertex in
    this set. A $k$-clique relaxes this requirement, requiring vertices to be within a distance $k$
    of each other, rather than directly adjacent. In theory, a maximum clique algorithm can easily
    be adapted to solve the maximum $k$-clique problem, although large sparse $k$-clique graphs
    reduce to large dense clique graphs, which can be computationally challenging. We adapt a state of
    the art maximum clique algorithm to show that this reduction is in fact useful in practice, and
    introduce a lazy global domination rule which sometimes vastly reduces the search space. We
    include experimental results for a range of real-world and benchmark graphs, and a detailed look
    at random graphs.
\end{abstract}

@@

A clique in a graph is a set of vertices, each of which is adjacent to every other vertex in the
set. Finding a clique of maximum size in a graph is one of the basic \NP-hard problems
\cite{Garey:1990}; applications include geometry, coding theory, computer vision and bioinformatics
\cite{Bomze:1999,Butenko:2006}. However, when analysing real-world data, a clique may be too strong
a requirement. A $k$-clique\footnote{Or sometimes $n$-clique or $s$-clique; in an unfortunate clash
of notation, ``$k$-clique problem'' is sometimes used elsewhere for the decision version of the
clique problem, to distinguish it from the maximum clique problem.} is a relaxed form of clique,
where instead of requiring each pair of vertices to be directly adjacent, we only require they be
connected by a path of length at most $k$ \cite{Luce:1950}.  Thus a $1$-clique \emph{is} a clique, a
$2$-clique may be thought of as ``a group of people, all of whom either know each other or have a
mutual acquaintance'', and so on. We illustrate this in \cref{figure:k-cliques}.  Determining the
size of a maximum $k$-clique is \NP-hard for any fixed $k$ \cite{Bourjolly:2002}.

@@

Here we give experimental results on a range of standard benchmarks, and on real-world and random
graphs. Where timing results are reported, the experiments were run on a machine with dual Intel
E5-2687W v3 processors; single-threaded runtimes are given, except in \cref{table:parallel}, where
20 threads were used (our machine does not have hyper-threading enabled). The time taken to read in
the graph from a file is excluded, but preprocessing time (including the construction of $G^k$ and
the bitset encoding) is included. We use the term \emph{nodes} to refer to the number of recursive
calls made by the branch-and-bound part of the algorithm.

@@

\subsection{Parallel Search}

\begin{table}
    \tiny\setlength{\tabcolsep}{4pt}
    \begin{tabular}{l c rr rr rr}
        \toprule
        & & & & \multicolumn{2}{c}{Sequential} & \multicolumn{2}{c}{Parallel} \\
    \cmidrule(lr){5-6}
    \cmidrule(lr){7-8}
    Instance & \multicolumn{1}{c}{$k$} & \multicolumn{1}{c}{$D$} & \multicolumn{1}{c}{$\tilde{\omega}_k$} &
    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} &
    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} \\
    \midrule
    \inputhaxx{gen-table-parallel}
    \bottomrule
\end{tabular}
\caption{Experimental results for multi-threaded search on harder instances, using 20 threads. For
    each graph, we consider $k$ equal to 2, 3 and 4. In each case we show the
    density of $G^k$, the size of a maximum $k$-clique, and then for both the
    sequential algorithm and the parallel algorithm (with lazy global domination in both cases),
    the number of nodes required, and the runtime in seconds.}\label{table:parallel}
\end{table}
